Estimating Quantile Regressions with Multiple Fixed Effects through Method of Moments

### 2024-09-10 article/alt_article.qmd
Update Paper
WIth Comments from GJ

keywords: 
  - Fixed effects
  - Linear heteroskedasticity
  - Location-scale model
  - Quantile regression

date: last-modified
bibliography: bibliography.bib
format:
  pdf: 
    toc-depth: 2
    number-sections: true
    documentclass: article
    margin-bottom: 1in
    margin-top: 1in
    margin-left: 1in
    margin-right: 1in
    thanks: |
      The opinions expressed in this paper are those of the authors and do not necessarily reflect the views of the World Bank, its Board of Directors, or the countries it represents.

@@

    fontsize: 11pt
    csl: "templates/apa-no-ampersand.csl"
    template-partials: 
      - "templates/title.tex"
      - "templates/before-body.tex"
      - "templates/_authors.tex"      
---

<!--
How to Render this?
Lik this:
quarto render article.qmd --to elsevier-pdf

-->
{{< pagebreak >}}

@@

::: {#tbl-sim1}

@@

::: {#tbl-sim2}

@@

::: {#tbl-sim3}

@@

# Illustrative application

@@

|          | polityt |  lyp   | trade | p1564 |  p65   |  lspl  | oil_im | oil_ex |  ygap  |
|----------|:-------:|:------:|:-----:|:-----:|:------:|:------:|:------:|:------:|:------:|
| Location |         |        |       |       |        |        |        |        |        |
|  coeff   |  0.116  | -0.715 | 0.030 | 0.121 | 0.028  | 0.691  | -0.047 | -0.006 | 0.010  |
|  se_gls  |  0.046  | 0.540  | 0.008 | 0.033 | 0.070  | 0.035  | 0.008  | 0.022  | 0.028  |
|  se_r    |  0.047  | 0.597  | 0.008 | 0.031 | 0.070  | 0.037  | 0.007  | 0.017  | 0.021  |
|  se_cl   |  0.046  | 0.465  | 0.007 | 0.032 | 0.071  | 0.035  | 0.010  | 0.020  | 0.023  |
| Scale    |         |        |       |       |        |        |        |        |        |
|  coeff   | -0.097  | -0.616 | 0.003 | 0.036 | 0.087  | -0.085 | 0.013  | 0.016  | -0.004 |
|  se_gls  |  0.032  | 0.371  | 0.005 | 0.023 | 0.048  | 0.024  | 0.006  | 0.015  | 0.019  |
|  se_r    |  0.031  | 0.398  | 0.005 | 0.020 | 0.049  | 0.025  | 0.005  | 0.010  | 0.015  |
|  se_cl   |  0.048  | 0.800  | 0.008 | 0.031 | 0.067  | 0.029  | 0.004  | 0.010  | 0.012  |
| Q25      |         |        |       |       |        |        |        |        |        |
|  coeff   |  0.191  | -0.239 | 0.028 | 0.093 | -0.039 | 0.756  | -0.057 | -0.018 | 0.013  |
|  se_gls  |  0.059  | 0.684  | 0.010 | 0.042 | 0.088  | 0.045  | 0.010  | 0.027  | 0.035  |
|  se_r    |  0.056  | 0.656  | 0.008 | 0.036 | 0.086  | 0.040  | 0.010  | 0.020  | 0.025  |
|  se_cl   |  0.073  | 0.687  | 0.006 | 0.041 | 0.098  | 0.023  | 0.010  | 0.021  | 0.029  |
| Q50      |         |        |       |       |        |        |        |        |        |
|  coeff   |  0.108  | -0.765 | 0.030 | 0.124 | 0.035  | 0.684  | -0.046 | -0.005 | 0.009  |
|  se_gls  |  0.046  | 0.535  | 0.007 | 0.033 | 0.069  | 0.035  | 0.008  | 0.022  | 0.027  |
|  se_r    |  0.046  | 0.593  | 0.008 | 0.031 | 0.069  | 0.036  | 0.007  | 0.017  | 0.021  |
|  se_cl   |  0.043  | 0.484  | 0.008 | 0.032 | 0.070  | 0.036  | 0.010  | 0.020  | 0.023  |
| Q75      |         |        |       |       |        |        |        |        |        |
|  coeff   |  0.031  | -1.258 | 0.033 | 0.153 | 0.104  | 0.616  | -0.036 | 0.008  | 0.006  |
|  se_gls  |  0.048  | 0.551  | 0.008 | 0.034 | 0.071  | 0.036  | 0.008  | 0.022  | 0.028  |
|  se_r    |  0.049  | 0.696  | 0.009 | 0.034 | 0.075  | 0.043  | 0.007  | 0.018  | 0.023  |
|  se_cl   |  0.039  | 0.919  | 0.012 | 0.041 | 0.079  | 0.055  | 0.010  | 0.022  | 0.020  |

: Determinants of Government Surpluses: Individual Fixed Effects {#tbl-tb1}

@@

|          | polity |  lyp   | trade | prop1564 | prop65 |  lspl  |  ygap  |
|----------|:------:|:------:|:-----:|:--------:|:------:|:------:|:------:|
| Location |        |        |       |          |        |        |        |
|  coeff   | 0.126  | -0.418 | 0.028 |  0.108   | 0.042  | 0.693  | -0.014 |
|  se_gls  | 0.087  | 1.157  | 0.015 |  0.072   | 0.136  | 0.066  | 0.053  |
|  se_r    | 0.047  | 0.703  | 0.008 |  0.038   | 0.068  | 0.038  | 0.022  |
|  se_cl   | 0.048  | 0.506  | 0.008 |  0.044   | 0.077  | 0.037  | 0.022  |
| Scale    |        |        |       |          |        |        |        |
|  coeff   | -0.095 | -1.255 | 0.005 |  0.033   | 0.040  | -0.081 | 0.008  |
|  se_gls  | 0.081  | 1.073  | 0.014 |  0.067   | 0.126  | 0.061  | 0.049  |
|  se_r    | 0.031  | 0.452  | 0.005 |  0.025   | 0.045  | 0.025  | 0.017  |
|  se_cl   | 0.041  | 0.848  | 0.006 |  0.030   | 0.048  | 0.033  | 0.013  |
| Q25      |        |        |       |          |        |        |        |
|  coeff   | 0.201  | 0.576  | 0.025 |  0.082   | 0.010  | 0.757  | -0.020 |
|  se_gls  | 0.154  | 2.070  | 0.024 |  0.118   | 0.219  | 0.121  | 0.085  |
|  se_r    | 0.058  | 0.751  | 0.008 |  0.049   | 0.080  | 0.040  | 0.026  |
|  se_cl   | 0.073  | 0.761  | 0.006 |  0.052   | 0.087  | 0.023  | 0.027  |
| Q50      |        |        |       |          |        |        |        |
|  coeff   | 0.119  | -0.512 | 0.029 |  0.111   | 0.045  | 0.687  | -0.013 |
|  se_gls  | 0.091  | 1.230  | 0.014 |  0.070   | 0.130  | 0.072  | 0.051  |
|  se_r    | 0.046  | 0.695  | 0.008 |  0.037   | 0.068  | 0.038  | 0.022  |
|  se_cl   | 0.045  | 0.529  | 0.008 |  0.044   | 0.077  | 0.039  | 0.021  |
| Q75      |        |        |       |          |        |        |        |
|  coeff   | 0.041  | -1.555 | 0.033 |  0.138   | 0.078  | 0.619  | -0.007 |
|  se_gls  | 0.067  | 0.898  | 0.011 |  0.053   | 0.098  | 0.052  | 0.038  |
|  se_r    | 0.048  | 0.827  | 0.009 |  0.037   | 0.075  | 0.046  | 0.026  |
|  se_cl   | 0.038  | 0.980  | 0.012 |  0.050   | 0.086  | 0.063  | 0.020  |

: The Determinants of Government Surpluses: Individual and Time Fixed Effects {#tbl-tb2}

@@

@@

>>>>

thanks: |
  The opinions expressed in this paper are those of the authors and do not necessarily reflect the views of the World Bank, its Board of Directors, or the countries it represents.

@@

keywords: [Fixed effects, Linear heteroskedasticity, Location-scale model, Quantile regression]
jelcode: [C21, C22, C23]

date: last-modified
bibliography: bibliography.bib
format:
  sim-art2-pdf:
    keep-tex: true

---

@@

::: {#tbl-sim1 tbl-pos="H"}

@@

::: {#tbl-sim2 tbl-pos="H"}

@@

::: {#tbl-sim3 tbl-pos="H"}

@@

# Illustrative applications

## Determinants of government surpluses

@@

::: {#tbl-tb1 tbl-pos="h"}

{{< include tbl4.tex >}}

Determinants of Government Surpluses: Individual Fixed Effects

:::

@@

:::{#tbl-tb2 tbl-pos="h"}

{{< include tbl5.tex >}}

The Determinants of Government Surpluses: Individual and Time Fixed Effects

:::

## The Effect of Criminal Activities on Labor Incomes{#sec-crime}

In this section, we investigate the effects of criminal activities on labor incomes using a quantile regression approach to discover whether crime has different impacts on the distribution of interest. We use data from the ENEMDU labor survey from Ecuador and combine it with administrative records of crime reports by Ecuadorian Cantons.

The combined data set covers the period 2019-2022. For the sake of brevity, we will only present here the results of quantile regressions of the log of total labor income on the log of annual reported murders (LMURDERS) and other covariates displayed in Table 1. However, our findings are robust to alternative specifications of the model where the dependent variable is either the log of per-capita household income or the log of total labor income from main occupation, both of which are displayed in @sec-app-crime.

:::{#tbl-ecuador tbl-pos="htbp"}

{{< include tbl6.tex >}}

The Effect of Crime on Labor Income: Canton, Activity Sector and Time Fixed Effects

:::

The findings in @tbl-ecuador support the idea that the effect of murders diminishes as we move forward in the distribution from low to high incomes: in the lowest decile, a 1% increase in reported murders is associated with a 3.08% reduction in total labor income; instead, for the top 10% labor income earners, murders are seen to have little effect on the dependent variable and, in addition, the coefficient is not significant at conventional levels. Quantile regression estimates are thus really useful to discover relationships between dependent and explanatory variables that are not seen in standard (mean) regression analysis. For this application, we have found a monotonic decreasing effect of the log of reported crimes on total labor income.

While the capability of estimating distinct coefficients of explanatory variables for each quantile is a feature present since @koenkerbasset1978, the method of moments approach allows us to include not one but multiple sets of fixed effects so that one can account for unobserved heterogeneity at different levels. Regarding this application, we have included in the estimation fixed effects for Canton, Activity Sector of the income earner and Year as is shown in @tbl-ecuador. Given that LMURDERS is measured at the Canton level, it may be important to control for unobserved factors affecting both crime and income earnings in the Cantons of Ecuador.

Finally, our approach also allows us to obtain clustered standard errors for estimated coefficients so that statistical inference can be performed properly. @tbl-ecuador presents clustered standard errors at the Canton level, the rationale being the same as described above for the inclusion of Canton fixed effects in the specification. Had we not taken into account the correlated nature of our standard errors, we would have computed misleading (and smaller) estimates. For example, robust standard errors in this specification are, on average, 2.5 times smaller than clustered standard errors. It is important, however, to remark that the effects of murders on labor income at the lower half of the distribution are still statistically different from zero when clustering at the Canton level.

@@

Our example using data to analyze the impact of Crime on labor income in Ecuador also suggests the relevance of using clustered standard errors in the presence of unobserved heterogeneity at different levels. We find that the effect of murders on labor income diminishes as we move forward in the distribution from low to high incomes, which is a result that would not have been possible without the use of quantile regression.

@@

# Complementary analysis of @sec-crime{#sec-app-crime}

In this appendix, we present the quantile regressions of both (log) per-capita household income and the (log) of labor income from main occupation on the (log) of annual reported murders. While in the first specification the coefficients of interest are not significant at conventional levels because of clustered standard errors (see @tbl-acrime1), the same pattern of decreasing association between murders and measurements of well-being as we move to the right of the distribution is revealed using our proposed method.

However, in @tbl-acrime2 we get significant (at the 1% level) associations between reported murders and a slightly different measurement of labor income that only takes into account the main occupation of the individual. Compared to total labor income, the coefficients displayed in @tbl-acrime2 are smaller in magnitude: for the lowest decile of the distribution, we find a 1 percentage point difference in the coefficient of interest between the two alternative specifications, where the largest effect corresponds to total labor income. Nevertheless, and once again, the decreasing pattern of reported murders across quantiles is documented for this specification. What is even more interesting, compared to our estimations in @tbl-ecuador, is that crime is no longer statistically meaningful from the median of the distribution onwards.

:::{#tbl-acrime1 tbl-pos="H"}

{{< include tbla1.tex >}}

The Effect of Crime on Per Capita Household Income: Canton and Time Fixed Effects
:::

:::{#tbl-acrime2 tbl-pos="H"}

{{< include tbla2.tex >}}

The Effect of Crime on Labor Income in Main Occupation: Canton, Activity
Sector and Time FE
:::
